\ourlongtitle

### 2021-08-24 Bernoulli/glo-2020-numerics.tex
corrected typos (Bernoulli revision)

In this toy example, our aim is to illustrate Theorem~\ref{thm:bias:bound} by viewing models with $\precpar > 0$ as `skew' versions of a `true' model with $\precpar = 0$. Given simulated data $(y_n)_{n \in \nsetpos}$ from the true model, smoothed additive expectations under skew models with different $\precpar > 0$ will be compared to the same expectations under the true model. Note that the model specified above satisfies condition \hypref{assum:bias:bound}; indeed, denote by $\md{}^\precpar(x_n, y_n)$ the emission density of $Y_n$ given $X_n$, which is Gaussian with mean $(1 - \precpar) \varphi(x_n)$ and unit variance, and by $\md{}(x_n, y_n)$ the same density for the true model. Then by the mean-value theorem, for all $\precpar \in \precparsp$ and $(x_n, y_n) \in \rset^2$, 
$$
\left\vert \md{}^\precpar(x_n, y_n) - \md{}(x_n, y_n) \right \vert \leq \precpar \sup_{\precpar \in \precparsp} \left \vert\frac{\partial}{\partial \precpar} \md{}^\precpar(x_n, y_n) \right\vert 
\leq \varepsilon \left( \vert \varphi(x_n) y_n - \varphi^2(x_n) \vert + \varphi^2(x_n) \right), 
$$
implying immediately that for all bounded measurable real-valued functions $h$,  
\begin{align*}
\left \vert \ukmod{n} h(x_n) - \uk{n} h(x_n) \right \vert \leq \precpar c(y_{n + 1}) \| h\|_\infty, 
\end{align*}
where $\ukmod{n}(x_n, \rmd x_{n + 1}) \eqdef \hk_\delta(x_n, \rmd x_{n + 1}) \md{}^\precpar(x_{n + 1}, y_{n + 1})$ and $c(y_{n + 1}) \eqdef \| \varphi \|_\infty |y_{n + 1}| + 2 \| \varphi \|_\infty^2$. Thus, \hypref{assum:bias:bound} holds true under the mild assumption that $\sup_{n \in \nsetpos} |y_n| < \infty$.

@@

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{%OL-version/figures/
        vary_epsilon_fixed_n.png}
        \caption{}
        \label{fig:varying:precpar}
    \end{subfigure}
    ~ 
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{%OL-version/figures/
        fixed_epsilon_vary_n.png}
        \caption{}
        \label{fig:varying:n}
    \end{subfigure}
    \caption{(a) The red line is the deviation of $\postmod{0:n} h_n$ from $\post{0:n} h_n$ for varying $\precpar \in \{0, 0.05, 0.1, \ldots, 0.5 \}$ and fixed $n = 50$, computed by means of Kalman smoothing. Turquoise markers are similar errors obtained on the basis of $60$ replicates of the ideal PaRIS in Algorithm~\ref{alg:ideal:PaRIS} with $\N = 200$ particles and $\K = 2$ backward samples. (b) The red line is the same deviation for increasing $n \in \intvect{1}{50}$ and fixed $\precpar = 0.1$. Then green and blue lines correspond to errors of $60$ independent ideal PaRIS replicates obtained under the skew and true models, respectively, for $\N = 200$ and $\K = 2$.} \label{fig}
\end{figure}

>>>>

In this toy example, our aim is to illustrate Theorem~\ref{thm:bias:bound} by viewing models with $\precpar > 0$ as `skew' versions of a `true' model with $\precpar = 0$. Given simulated data $(y_n)_{n \in \nsetpos}$ from the true model, smoothed additive expectations under skew models with different $\precpar > 0$ will be compared to the same expectations under the true model. Note that the model specified above satisfies condition \hypref{assum:bias:bound}; indeed, denote by $\md{}^\precpar(x_n, y_n)$ the emission density of $Y_n$ given $X_n$, which is Gaussian with mean $(1 - \precpar) \varphi(x_n)$ and unit variance, and by $\md{}(x_n, y_n)$ the same density for the true model. Then by the mean-value theorem, for all $\precpar \in \precparsp$ and $(x_n, y_n) \in \rset^2$, 
$$
\left\vert \md{}^\precpar(x_n, y_n) - \md{}(x_n, y_n) \right \vert \leq \precpar \sup_{\precpar \in \precparsp} \left \vert\frac{\partial}{\partial \precpar} \md{}^\precpar(x_n, y_n) \right\vert 
\leq \varepsilon \left( \vert \varphi(x_n) y_n - \varphi^2(x_n) \vert + \varphi^2(x_n) \right), 
$$
so that for all bounded measurable real-valued functions $h$,  $\vert \ukmod{n} h(x_n) - \uk{n} h(x_n) \vert \leq \precpar c(y_{n + 1}) \| h\|_\infty$, where $\ukmod{n}(x_n, \rmd x_{n + 1}) \eqdef \hk_\delta(x_n, \rmd x_{n + 1}) \md{}^\precpar(x_{n + 1}, y_{n + 1})$ and $c(y_{n + 1}) \eqdef \| \varphi \|_\infty |y_{n + 1}| + 2 \| \varphi \|_\infty^2$. Thus, \hypref{assum:bias:bound} holds true under the mild assumption that $\sup_{n \in \nsetpos} |y_n| < \infty$.

@@

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{%OL-version/figures/
        vary_epsilon_fixed_n.png}
        \caption{}
        \label{fig:varying:precpar}
    \end{subfigure}
    ~ 
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{%OL-version/figures/
        fixed_epsilon_vary_n.png}
        \caption{}
        \label{fig:varying:n}
    \end{subfigure}
    \caption{(a) The red line is the deviation of $\postmod{0:n} h_n$ from $\post{0:n} h_n$ for varying $\precpar \in \{0, 0.05, 0.1, \ldots, 0.5 \}$ and fixed $n = 50$, computed by means of Kalman smoothing. Turquoise markers are similar errors obtained on the basis of $60$ replicates of the ideal PaRIS in Algorithm~\ref{alg:ideal:PaRIS} with $\N = 200$ particles and $\K = 2$ backward samples. (b) The red line is the same deviation for increasing $n \in \intvect{1}{50}$ and fixed $\precpar = 0.1$. Then green and blue lines correspond to errors of $60$ independent ideal PaRIS replicates obtained under the skew and true models, respectively, for $\N = 200$ and $\K = 2$.} \label{fig}
\end{figure}
